Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-10-31 ms_pre_post.Rmd
Edit Model: Mating section

@@

Each male within the population is a potential mate for any female to choose. We therefore assume that there are no opportunity costs associated with male mating such that mating with one female would reduce a male's ability to mate with and sire the offspring of any other female [e.g., @Waser1986]. While all males are therefore available as potential mates, we assume that a female mates with her total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models the total number of distinct males that a female chooses to mate with rather than the number of times that she mates. In the unlikely event that a female's $N_{males}$ exceeds the total number of males in the population, then she mates with all of the males in the population. Most often, $N_{males}$ will be a small subset of the available males in the population [@Duthie], so the male(s) a focal female chooses will reflect her expression of pre-copulatory inbreeding strategy phenotype ($M_{p}$). Like polyandry allele values, pre-copulatory inbreeding strategy allele values ($M_{a}$) can take any real number. However, unlike polyandry, the sum of all of a female's 20 $M_{a}$ alleles equals both her pre-copulatory inbreeding strategy genotype ($M_{g}$) and her pre-copulatory inbreeding strategy phenotype ($M_{p}$). Negative or positive values of $M_{p}$ cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ cause females to mate randomly with respect to kinship.

The probability that a focal female $i$ mates with a male $j$ with which she shares some kinship $k_{ij}$ is calculated by first assigning each male a perceived mate quality $Q^{m}_{ij}$. This quality depends upon whether the female has a strategy of avoiding mating with kin ($M_{p}<0$) or preferring mating with kin ($M_{p}>0$; if $M_{p}=0$, all males are assigned a quality of 1, $Q^{m}_{ij}=1$). If a focal female avoids mating with kin, then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that the perceived quality of a male decreases linearly with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If a focal female prefers mating with kin, then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that the perceived quality of a male increases with increasingly positive values of $k_{ij}$ and $M_{p}$. After a focal female assigns all males a perceived quality, the sum of all male perceived qualities is calculated, then each male quality is divided by this sum. This transformation results in each male having a relative perceived mate quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a focal female chooses the male as a mate; female choice is therefore stochastic, meaning that females do not always choose the male of the highest $q^{m}_{ij}$ (e.g., given $M_{p}$ values of $-2$, $-4$, $-6$, and $-8$, an outbred focal female is $1.5$, $2$, $2.5$, and $3$ times more likely to mate with a non-relative than a full-sibling $k_{ij}=0.25$, respectively). For polyandrous females that choose more than one male mate, mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement.

>>>>

All males in the population are assumed to be available for any female to choose. We therefore assume that there is no opportunity cost of male mating such that mating with one female would reduce a male's availability to mate with any other female. Individual females mate with their total allotment of $N_{males}$ without replacement, meaning that $N_{males}$ models a female's total number of different mates rather than solely her total number of matings.

Most often, $N_{males}$ will be smaller than the total number of available males [@Duthie]. Here, each female chooses her $N_{males}$ mates based on her pre-copulatory inbreeding strategy phenotype ($M_{p}$), which equals her pre-copulatory inbreeding strategy genotype ($M_{g}$) defined as the sum of the values of her 20 $M_{a}$ alleles. Negative or positive $M_{p}$ values cause a female to avoid or prefer mating with kin, respectively. Values of $M_{p}=0$ cause females to mate randomly with respect to kinship.

The probability that a focal female $i$ mates with a male $j$ with which she shares some kinship $k_{ij}$ is calculated by first assigning each male a perceived mate quality $Q^{m}_{ij}$. If the female has a strategy of pre-copulatory inbreeding avoidance ($M_{p}<0$), then $Q^{m}_{ij} = (-M_{p} \times k_{ij} + 1)^{-1}$, meaning that $Q^{m}_{ij}$ decreases linearly with increasingly positive values of $k_{ij}$ and increasingly negative values of $M_{p}$. If the female has a strategy of pre-copulatory inbreeding preference ($M_{p}>0$), then $Q^{m}_{ij} = M_{p} \times k_{ij} + 1$, meaning that $Q^{m}_{ij}$ increases with increasingly positive $k_{ij}$ and $M_{p}$. If $M_{p}=0$, then all males are assigned a quality of 1

After a focal female assigns all males a $Q^{m}_{ij}$ value, each male's value is divided by the sum of all $Q^{m}_{ij}$ values across all males, thereby assigning each male a relative perceived quality $q^{m}_{ij}$, which is constrained to values between zero and one. The value of $q^{m}_{ij}$ then defines the probability that a focal female mates with the male; mating is therefore stochastic, and females do not always mate with the male of the highest $q^{m}_{ij}$. For polyandrous females that choose multiple mates (i.e., $N_{mates}>1$), mates are chosen iteratively such that $Q^{m}_{ij}$ and $q^{m}_{ij}$ are re-calculated for each additional mate choice, and with $Q^{m}_{ij}$ and therefore $q^{m}_{ij}$ values of already chosen males set to zero to ensure mate sampling without replacement. In the unlikely event that a female's $N_{males}$ exceeds the total number of available males, then she simply mates with all males.

@@
